---
title: "hyperbandr tutorial"
author: "Niklas"
date: "23 März 2018"
output: 
  pdf_document: 
    keep_tex: yes
header-includes:
  - \usepackage{graphicx}
---

# C:\Users\Niklas\AppData\Local\Temp

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is an R6 implementation of the original **hyperband** algorithm  <https://arxiv.org/abs/1603.06560>.

R6 is an encapsulated object oriented system akin to those in Java or C++, where objects contain methods in addition to data, and those methods can modify objects directly (unlike S3 and S4 which are both functional object-oriented systems, where class methods are separate from objects, and objects are not mutable).

Essentially, that means that we obtain a very generic implementation, which is working with every other R package (as long the algorithm meets the requirements of hyperband).

## This tutorial consists of four examples:

* hyperband to optimize a **neural network** with *mxnet* and *mlr*
* hyperband in combination with MBO to optimize a **neural network** with *mxnet*, *mlr* and *mlrMBO*
* hyperband to optimize a **boosting** model with *xgboost* and *mlr*
* hyperband to optimize a **function** with *smoof*

\newpage
## Example 1: hyperband to optimize a **neural network** with *mxnet* and *mlr*

```{r, echo = TRUE, message = FALSE}
library("R6")
library("devtools")
load_all()
library("mxnet")
library("mlr") # you might need to install mxnet branch of mlr: devtools::install_github("mlr-org/mlr", ref = "mxnet")
library("ggplot2")
library("gridExtra")
library("dplyr")
library("data.table")
# read mini_mnist (1/10 of actual mnist for faster evaluation, evenly distributed classes)
train = fread("../mnist/train.csv", header = TRUE)
test = fread("../mnist/test.csv", header = TRUE)

# Some operations to normalize features
mnist = as.data.frame(rbind(train, test))
mnist = mnist[sample(nrow(mnist)), ]
mnist[, 2:785] = lapply(mnist[, 2:785], function(x) x/255)
rm(train)
rm(test)
```

```{r, echo = FALSE}

```



```{r, echo = TRUE}
dim(mnist)
```



```{r, echo = TRUE}

```



```{r, echo = TRUE}

```


